---
title: "TFM"
author: "Karolina Gliwa"
date: "2024-02-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE )
```

```{r data,eval=FALSE}

#NASA
######################### OCEAN WARMING ######################

url <- "https://www.ncei.noaa.gov/data/oceans/woa/DATA_ANALYSIS/3M_HEAT_CONTENT/DATA/basin/pentad/pent_h22-w0-2000m.dat"
ocean_warming <- read.table(url, header = TRUE, sep = "", fill = TRUE, check.names = TRUE)

str(data)
########################## CARBON #########################
urll="https://gml.noaa.gov/webdata/ccgg/trends/co2/co2_mm_mlo.txt"
carbon <- read.table(urll, header = TRUE, sep = "", fill = TRUE, check.names = TRUE)

str(carbon)

colnames(carbon) <- c("Year", "Month", "decimal_date", "monthly_average", "de_seasonalized", "days", "st_dev", "unc_of_mon_mean")

##################### TEMPERATURES ####################
urlll="https://data.giss.nasa.gov/gistemp/graphs/graph_data/Global_Mean_Estimates_based_on_Land_and_Ocean_Data/graph.txt"
temperatures <- read.table(urlll, header = TRUE, sep = "", fill = TRUE, check.names = TRUE)

str(temperatures)

temperatures$X.C.=NULL
# Elimina las tres primeras filas
temperatures <- temperatures[-c(1:3), ]

# Renombra las columnas según la segunda imagen proporcionada
colnames(temperatures) <- c("Year", "No_Smoothing", "Lowess_5")
str(temperatures)

temperatures$Year=as.numeric(temperatures$Year)
temperatures$No_Smoothing=as.numeric(temperatures$No_Smoothing)
temperatures$Lowess_5=as.numeric(temperatures$Lowess_5)

################# SEA LEVEL ##################################

urlllll="https://deotb6e7tfubr.cloudfront.net/s3-edaf5da92e0ce48fb61175c28b67e95d/podaac-ops-cumulus-protected.s3.us-west-2.amazonaws.com/MERGED_TP_J1_OSTM_OST_GMSL_ASCII_V51/GMSL_TPJAOS_5.1.txt?A-userid=karolinagliwa&Expires=1711726958&Signature=E~ZJ68OfPZV6Oo32VolCtXWfg-2T1m-y4uUMqwxAwVes-B8RHi2ECI1~FXmF8S5uaWS7wYLC8It7p44OFHBOr~Fof74PRxHS~9HY9bGYLdjjZT14LT-UeSzxfWX6HL0pm3A6mX4fg0N3Zb4cDlRDnwZP5emqWnhERF8gZk8mGvglUVXgN8-yVpN4iPxjUG8hBkzdCruBJaZxmfO3ztCcHaLgH3n7tj0foW4uPXkc0CnyYybq64dwzCQH3NicpWFy5yiQv5O3hsBPS8KocQb4t8lVWc3d6rV4ZjR5NSn9LYY1kMJgcm8Txg3gd~KrNFUGafHk~s2OJq5ZXL5msP-yVg__&Key-Pair-Id=KZRIRLMTAHKYV"
header_lines <- 52

# Read the data, skipping the header rows
sea_level <- read.table(urlllll, skip = header_lines, header = FALSE,fill = TRUE, sep = "")

# Manually assign column names based on your description
colnames(sea_level) <- c("Altimeter_Type", "Merged_File_Cycle_Number", "Year_Fraction",
                         "Number_of_Observations", "Number_of_Weighted_Observations",
                         "GMSL_No_GIA_Variation", "GMSL_No_GIA_SD",
                         "Smoothed_GMSL_No_GIA", "GMSL_GIA_Variation",
                         "GMSL_GIA_SD", "Smoothed_GMSL_GIA",
                         "Smoothed_GMSL_GIA_Annual_Semiannual_Removed",
                         "Smoothed_GMSL_No_GIA_Annual_Semiannual_Removed")

# View the first few rows of the data to confirm it's read correctly
#head(gmsl_data)

######################## METHANE ###########################
urllll="https://gml.noaa.gov/webdata/ccgg/trends/ch4/ch4_annmean_gl.txt"
methane <- read.table(urllll, header = TRUE, sep = "", fill = TRUE, check.names = TRUE)
str(methane)
colnames(methane)=c("year","mean","unc")

######################### ARCTIC SEA #########################
Arctic_sea_ice=c(6.14,6.47,7.47,6.4,7.14,6.08,7.58,6.69,6.54,6.12,6.25,6.73,5.83,6.12,5.98,5.5,5.86,4.27,4.69,5.26,4.87,4.56,3.57,5.21,5.22,4.62,4.53,4.82,4.79,4.36,4,4.95,4.9)


# WORLD_BANK

refugees=readxl::read_excel("refugees.xlsx")
str(refugees)
refugees$year <- gsub("\\s*\\[.*?\\]\\s*", "", refugees$year)
refugees$year <- trimws(refugees$year)
str(refugees)
refugees$year=as.numeric(refugees$year)
refugees$`Refugee population`=as.numeric(refugees$`Refugee population`)
colnames(refugees)=c("year","refugees")



#disasters=read.csv("disasters.csv")


# IMF DISASTERS (count world):

Drought=c(19,12,11,9,11,6,7,17,20,26,25,22,24,14,11,20,9,11,16,18,21,16,18,9,20,27,14,12,17,15,12,16,22)

#Extreme_temperature=c(13,8,7,3,9,13,5,13,12,8,30,23,15,25,16,40,14,25,16,20,28,15,52,13,17,12,12,10,27,21,5,3,12)

Flood=c(59,75,57,84,88,93,91,95,94,122,157,155,171,157,127,191,226,217,165,151,184,156,136,148,136,161,159,127,127,195,201,222,176)

Landslide=c(5,9,8,25,11,16,24,13,22,18,27,24,20,21,15,13,20,10,12,29,32,17,13,11,15,20,13,25,12,25,19,12,17)

Storm=c(132,66,76,105,80,76,74,79,86,103,101,105,122,85,121,130,76,100,112,87,94,81,90,105,98,121,86,126,94,91,127,119,105)

Wildfire=c(4,8,8,2,13,7,4,15,18,21,30,14,22,14,8,13,9,18,5,9,7,8,6,10,4,13,10,15,10,14,9,19,15)


```

# 1.INTRODUCTION

## 1.1 VARIABLES

-   World Bank Variables (UNHCR):
    -   **Refugees**: Forcibly displaced people worldwide
    -   **Remittances**: The money or goods that migrants send back to families and friends in origin countries
-   NASA Variables:
    -   **Carbon**: Monthly mean CO2 constructed from daily mean values.(CO2 parts per million/ppm)
    -   **Methane**: Mean per year of CH4 parts per billion/ ppb
    -   **Temperature**: Annual average Temperature Anomaly (C), change in global surface temperature compared to the long-term average from 1951 to 1980.
    -   **Ocean Warming**: Ocean warming measurement in Zettajoules, yearly data of world ocean warming with reference in 1955.
    -   **Sea Level**: Sea Height Variation (mm) from 1993
    -   **Arctic Sea Ice**: Arctic Sea Ice Minimum Extent in Million Square km
-   IMF Variables:
    -   **Climate-related Disasters Frequency (Count WorldWide)**:
        -   Drought
        -   Flood
        -   Landslide
        -   Storm
        -   Wildfire

```{r cleaning, eval=FALSE}

# REFUGEES
refugees=na.omit(refugees)
data=data.frame(refugees)

# CARBON
carbon=carbon[,c(1,2,4)]
carbon$Date <- as.Date(paste(carbon$Year, carbon$Month, "01", sep="-"))
carbon=carbon[carbon$Year >= 1990 & carbon$Year <=2022, ]
carbon = aggregate(monthly_average ~ Year, data = carbon, FUN = mean)
data$carbon=carbon$monthly_average

# METHANE
methane=methane[methane$year >= 1990, c(1,2) ]
data$methane=methane$mean

# TEMPERATURE
str(temperatures)
temperatures=temperatures[temperatures$Year >= 1990 & temperatures$Year <= 2022, ]
str(temperatures)
data$temperatures=temperatures$No_Smoothing

# OCEAN WARMING
ocean_warming=ocean_warming[ocean_warming$YEAR >= 1990.5, c(1,2)]
# this variable is until 2021 so we will need to predict the last year
str(ocean_warming)
new_row=data.frame(YEAR = 2022.5, WO = NA)
ocean_warming=rbind(ocean_warming, new_row)
data$ocean_warming=ocean_warming$WO



# SEA LEVEL
sea_level=sea_level[sea_level$Year_Fraction >= 1990 & sea_level$Year_Fraction < 2023, c(3,9)]
# this variable is until 2021 so we will need to predict the last year
str(sea_level)
# Convert 'Year_Fraction' to integer to extract the year part
#sea=floor(sea_level$Year_Fraction)
# Count the number of rows for each year
#year_counts= table(sea)
#print(year_counts)
# in most years we have 37 data points, in some 36, we do the mean of them to get the data yearly:
library(dplyr)

df=sea_level %>%
  mutate(Year = floor(Year_Fraction))

# Then group by the new Year column and summarize by taking the mean
yearly_means_sea <- df %>%
  group_by(Year) %>%
  summarize(Mean_GMSL_GIA_Variation = mean(GMSL_GIA_Variation))

# three years missing at the beginning so lets put NA:
new_row_sea=data.frame(Year = c(1990,1991,1992), Mean_GMSL_GIA_Variation = NA)
sea_level=rbind(yearly_means_sea, new_row_sea)
n=nrow(sea_level)
sea_level=sea_level[c((n-2):n, 1:(n-3)), ]
data$sea_level=sea_level$Mean_GMSL_GIA_Variation


# Arctic Sea Ice:
data$arctic_ice=Arctic_sea_ice


# DISASTERS:

data$Drought=Drought
#data$Extreme_Temperatures=Extreme_temperature
data$Flood=Flood
data$Landslide=Landslide
data$storm=Storm
data$wildfire=Wildfire



data$remittence=c(0.39656088,0.331430354,0.339577476,0.333704435,0.342751418,	0.315735097,0.318308727,0.364217975,0.361037066,0.363641584,0.371527886,0.40481814,	0.451069703,0.494541305,0.503232142,0.540250707,0.577167169,0.612575519,0.64748971,	0.654435339,0.641989972,0.649475952,0.666534956,0.688768824,0.71613182,0.749652787,	0.723788794,0.740057663,0.748774023,0.767368793,0.788959546,0.779994364,0.796443702)


write.csv(data, file.path(getwd(), "data.csv"))

```

```{r ts}

data <- read.csv("data.csv")
data=data[,-1]

#ts_refugees <- ts(data$refugees, start=c(min(data$year)), frequency=1)
ts_carbon <- ts(data$carbon, start=c(min(data$year)), frequency=1)
ts_methane <- ts(data$methane, start=c(min(data$year)), frequency=1)
ts_temperatures <- ts(data$temperatures, start=c(min(data$year)), frequency=1)
ts_ocean_warming <- ts(data$ocean_warming, start=c(min(data$year)), frequency=1)
ts_sea_level <- ts(data$sea_level, start=c(min(data$year)), frequency=1)
ts_drought <- ts(data$Drought, start=c(min(data$year)), frequency=1)
#ts_extreme_temperatures <- ts(data$Extreme_Temperatures, start=c(min(data$year)), frequency=1)
ts_flood <- ts(data$Flood, start=c(min(data$year)), frequency=1)
ts_landslide <- ts(data$Landslide, start=c(min(data$year)), frequency=1)
ts_storm <- ts(data$storm, start=c(min(data$year)), frequency=1)
ts_wildfire <- ts(data$wildfire, start=c(min(data$year)), frequency=1)
ts_arctic_ice <- ts(data$arctic_ice, start=c(min(data$year)), frequency=1)
ts_remittence <- ts(data$remittence, start=c(min(data$year)), frequency=1)


ts_list <- list(
  #refugees = ts_refugees,
  carbon = ts_carbon,
  methane = ts_methane,
  temperatures = ts_temperatures,
  ocean_warming = ts_ocean_warming,
  sea_level = ts_sea_level,
  drought = ts_drought,
  #extreme_temperatures = ts_extreme_temperatures,
  flood = ts_flood,
  landslide = ts_landslide,
  storm = ts_storm,
  wildfire = ts_wildfire,
  arctic_ice = ts_arctic_ice,
  remittence=ts_remittence
)

library(reticulate)

```

## 1.2 PLOTS

```{r}

library(reticulate)
use_python("C:/Users/invitado.pc/anaconda3/python.exe", required = TRUE)

```

```{python}

import pandas as pd
import statsmodels.api as sm

data = pd.read_csv("data.csv")
# Remove the first column from the DataFrame
data = data.iloc[:, 1:]

# Function to create a time series in Python
def create_time_series(column, start_year):
    return pd.Series(data[column].values, index=pd.date_range(start=f"{start_year}", periods=len(data), freq='A'))

# Determine the sprint(data.info())
start_year = data['year'].min()

# Creating time series for each metric
#ts_refugees = create_time_series('refugees', start_year)
ts_carbon = create_time_series('carbon', start_year)
ts_methane = create_time_series('methane', start_year)
ts_temperatures = create_time_series('temperatures', start_year)
ts_ocean_warming = create_time_series('ocean_warming', start_year)
ts_sea_level = create_time_series('sea_level', start_year)
ts_drought = create_time_series('Drought', start_year)
# ts_extreme_temperatures = create_time_series('Extreme_Temperatures', start_year)  # Uncomment if needed
ts_flood = create_time_series('Flood', start_year)
ts_landslide = create_time_series('Landslide', start_year)
ts_storm = create_time_series('storm', start_year)
ts_wildfire = create_time_series('wildfire', start_year)
ts_arctic_ice = create_time_series('arctic_ice', start_year)
ts_remittences= create_time_series('remittence', start_year)


ts_list = {
    #'refugees': ts_refugees,
    'carbon': ts_carbon,
    'methane': ts_methane,
    'temperatures': ts_temperatures,
    'ocean_warming': ts_ocean_warming,
    'sea_level': ts_sea_level,
    'drought': ts_drought,
    # 'extreme_temperatures': ts_extreme_temperatures,  # Uncomment if needed
    'flood': ts_flood,
    'landslide': ts_landslide,
    'storm': ts_storm,
    'wildfire': ts_wildfire,
    'arctic_ice': ts_arctic_ice,
    'remittances':ts_remittences
}
```

```{python}
import matplotlib.pyplot as plt
n = len(ts_list)  # Number of time series
cols = 3  # Number of columns in the subplot grid
rows = n // cols + (n % cols > 0)  # Calculate rows needed, rounding up

# Create a large figure to hold all subplots
plt.figure(figsize=(15, rows * 3))

# Loop through the dictionary and create a subplot for each time series
for i, (name, ts) in enumerate(ts_list.items(), start=1):
    plt.subplot(rows, cols, i)  # Create subplot
    ts.plot(title=name.capitalize())  # Plot time series
    plt.xlabel('Time')
    plt.ylabel(name.capitalize())

plt.tight_layout() # Adjust layout to prevent overlap

plt.savefig('time_series_grid.png')

plt.show() 

```

## 1.3 CHECK NAs:

```{r nas}
variables_with_na <- c() 
for(i in names(ts_list)) {
  if(any(is.na(ts_list[[i]]))) {  
    variables_with_na <- c(variables_with_na, i)  
  }
}

variables_with_na
 
```

### 1.3.1 Impute NAs

Package `imputeTS` for imputing NAs in time series:

The imputation method chosed is Kalman with Auto.Arima

The Kalman filter,it's a mathematical method that keeps updating and improving its estimates or predictions of what's happening in a process over time, even when some of the incoming information is incomplete or has some errors. It's constantly taking in new measurements, comparing them with its predictions, and then tweaking those predictions to be more accurate. Now, when it's paired with auto.arima, which picks the best ARIMA model we can put them to work together.

Auto.arima automatically finds patterns in the data---like trends and seasonal swings---and selects the best formula to describe these patterns. Then the Kalman filter uses this formula to make its "guesses" for the missing points.

```{r impnas, fig.align='center', warning=FALSE, message=FALSE}

library(gridExtra)
library(ggplot2)
library(ggtext)
library(stinepack)
library(imputeTS)

#plot_ocean_warming_before=ggplot_na_distribution(ts_ocean_warming)
plot_sea_level_before=ggplot_na_distribution(ts_sea_level,title = "Distribution of Missing Values Before Imputation")
#plot_remittance_before=ggplot_na_distribution(ts_remittence)


#ts_ocean_warming_imputed=na_kalman(ts_ocean_warming,model="auto.arima") 
ts_sea_level_imputed= na_kalman(ts_sea_level, model="auto.arima")
#ts_remittance_imputed= na_kalman(ts_remittence, model="auto.arima")

library(zoo)

ts_ocean_warming_imputed <- na.locf(ts_list$ocean_warming)
#ts_sea_level_imputed <- na.locf(ts_list$sea_level)
#ts_remittance_imputed <- na.locf(ts_list$remittence)


#plot_ocean_warming_after=ggplot_na_distribution(ts_ocean_warming_imputed)
plot_sea_level_after=ggplot_na_distribution(ts_sea_level_imputed,title = "Distribution of Missing Values After Imputation")
#plot_remitance_after=ggplot_na_distribution(ts_remittance_imputed)


#grid.arrange(plot_ocean_warming_before, plot_ocean_warming_after, ncol = 2, top="OCEAN WARMING")
grid_plot=grid.arrange(plot_sea_level_before, plot_sea_level_after, ncol = 2)
#grid.arrange(plot_remittance_before, plot_remitance_after, ncol = 2, top="REMITTENCE %")

ggsave("sea_level_grid.png", grid_plot, width = 10, height = 5,units = "in")

#ts_ocean_warming=ts_ocean_warming_imputed
ts_sea_level=ts_sea_level_imputed
#ts_remittence=ts_remittance_imputed



ts_list <- list(
 # refugees = ts_refugees,
  carbon = ts_carbon,
  methane = ts_methane,
  temperatures = ts_temperatures,
  ocean_warming = ts_ocean_warming,
  sea_level = ts_sea_level,
  drought = ts_drought,
  flood = ts_flood,
  landslide = ts_landslide,
  storm = ts_storm,
  wildfire = ts_wildfire,
  arctic_ice = ts_arctic_ice,
  remittence=ts_remittence
)



```

```{r}
variables_with_na <- c() 
for(i in names(ts_list)) {
  if(any(is.na(ts_list[[i]]))) {  
    variables_with_na <- c(variables_with_na, i)  
  }
}

variables_with_na
```

```{r}
ts_sea_level <- ts_list$sea_level

# Convert the R time series to a Python object
ts_sea_level_py <- r_to_py(ts_sea_level)
```

```{python}

ts_sea_level_py = r.ts_sea_level_py

ts_sea_level_series = pd.Series(ts_sea_level_py, index=pd.date_range(start='1990', periods=len(ts_sea_level_py), freq='A-DEC'))

ts_list = {
    #'refugees': ts_refugees,
    'carbon': ts_carbon,
    'methane': ts_methane,
    'temperatures': ts_temperatures,
    'ocean_warming': ts_ocean_warming,
    'sea_level': ts_sea_level_series,
    'drought': ts_drought,
    # 'extreme_temperatures': ts_extreme_temperatures,  
    'flood': ts_flood,
    'landslide': ts_landslide,
    'storm': ts_storm,
    'wildfire': ts_wildfire,
    'arctic_ice': ts_arctic_ice,
    'remittence':ts_remittences
}

```

As ocean warming has the last value as NA, we will eliminate it, so we will eliminate the last value of all time series

```{r}
ts_list=lapply(ts_list, function(ts) ts[-length(ts)])

ts_list=lapply(ts_list, function(ts) {
  ts(ts, frequency = 1, start = c(min(data$year)))})

```

```{python}

ts_list = {key: value[:-1] for key, value in ts_list.items()}

```

```{r}
variables_with_na <- c() 
for(i in names(ts_list)) {
  if(any(is.na(ts_list[[i]]))) {  
    variables_with_na <- c(variables_with_na, i)  
  }
}

variables_with_na
```

## 1.4 CORRELATIONS:

```{r corr, message=FALSE, warning=FALSE}

library(ggplot2)
library(forecast)
library(gridExtra)
library(tibble)
library(zoo)
library(tseries)
library(xtable)


ts_matrix <- do.call(cbind, lapply(ts_list, as.numeric))
cor_matrix <- cor(ts_matrix)
print(cor_matrix)
cor_matrix_1 <- cor_matrix[, 1:6]  # First 6 columns
cor_matrix_2 <- cor_matrix[, 7:12]  # Next 6 columns

# Convert each segment to a LaTeX table
cor_matrix_latex_1 <- xtable(cor_matrix_1, caption = "Correlation Matrix of Time Series (Part 1)")
cor_matrix_latex_2 <- xtable(cor_matrix_2, caption = "Correlation Matrix of Time Series (Part 2)")

# Save the LaTeX tables to .tex files
print(cor_matrix_latex_1, file = "correlation_matrix_1.tex")
print(cor_matrix_latex_2, file = "correlation_matrix_2.tex")
```

### 1.4.1 Multicollinearity:

```{r, warning=FALSE, message=FALSE}
library(car)

df <- data.frame(ts_list)

# Preliminary check for multicollinearity
# Fit a linear model with all variables as predictors
lm_model <- lm(remittence ~ ., data = df) 
vif_results <- vif(lm_model)
print(vif_results)


vif_df <- as.data.frame(vif_results)
vif_df$Variable <- rownames(vif_df)
rownames(vif_df) <- NULL

# Convert the VIF results to a LaTeX table
vif_latex <- xtable(vif_df, caption = "Variance Inflation Factor (VIF) Results")

# Save the LaTeX table to a .tex file
print(vif_latex, file = "vif_results.tex", include.rownames = FALSE)
```

### 1.4.2 Selection of Variables

We will eliminate one by one starting from the highest one.

```{r}

print("Carbon eliminated")
# carbon eliminated:
df_modified_vif_before=df[, !(colnames(df) %in% "carbon")]

lm_model <- lm(remittence ~ ., data = df_modified_vif_before) 
vif_results <- vif(lm_model)
print(vif_results)

print("------------------------------") # Separator


# sea_level eliminated:
print("Sea Level eliminated")
df_modified_vif_before=df_modified_vif_before[, !(colnames(df_modified_vif_before) %in% "sea_level")]
lm_model <- lm(remittence ~ ., data = df_modified_vif_before) 
vif_results <- vif(lm_model)
print(vif_results)

print("------------------------------") # Separator

# ocean_warming eliminated:
print("Ocean Warming eliminated")
df_modified_vif_before=df_modified_vif_before[, !(colnames(df_modified_vif_before) %in% "ocean_warming")]
lm_model <- lm(remittence ~ ., data = df_modified_vif_before) 
vif_results <- vif(lm_model)
print(vif_results)

```

Now all predictor variables have VIFs below 10. Multicollinearity solved.

# 2. ANALYSIS

## 2.1 STATIONARITY

We dont have seasonality as we have annual data, so we will check the trend:

-   If a series has a deterministic trend (e.g., it consistently increases by a certain amount each period), it is non-stationary because the mean changes over time. However, this non-stationarity can often be removed by detrending the series (e.g., by subtracting the fitted trend line from the original series).

-   If a series has a unit root, the non-stationarity cannot be removed simply by detrending because the underlying process has a stochastic trend. Instead, differencing the series (taking the change from one period to the next) is a common approach to achieve stationarity.

Unit root tests such as the Augmented Dickey-Fuller (ADF) test or the Phillips-Perron (PP) test are designed to distinguish between a unit root non-stationarity and other types of non-stationarity or stationarity. They test the null hypothesis that a unit root is present against the alternative that the series is stationary or has a deterministic trend.

The ADF test addresses whether a time series is non-stationary and possesses a unit root; its null hypothesis is that the series has a unit root (indicating non-stationarity), against the alternative hypothesis of stationarity. The test includes lagged difference terms in its regression to account for autocorrelation.

The PP test, which also tests for a unit root, differs from the ADF by considering the serial correlation in the error terms through non-parametric statistical adjustments, thus allowing for a broader test for non-stationarity without specifying a model for the autocorrelation structure.

### 2.1.1 Augmented Dickey-Fuller (ADF)

```{python}
import pandas as pd
import numpy as np
from arch.unitroot import *


# Function to perform the ADF test and return the t-statistic and p-value
def run_adf(ts, trend='c'):
    adf_test = ADF(ts.dropna(), trend=trend)
    return adf_test.stat, adf_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'

# Assuming ts_list is a dictionary with time series data as its values
# The keys of the dictionary are the names of the variables

# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_adf(ts, trend='c')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(t_stat_level)
    results_constant['p-value Level'].append(p_value_level)
    results_constant['t-value First Difference'].append(t_stat_diff)
    results_constant['p-value First Difference'].append(p_value_diff)
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_adf(ts, trend='ct')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(t_stat_level)
    results_trend['p-value Level'].append(p_value_level)
    results_trend['t-value First Difference'].append(t_stat_diff)
    results_trend['p-value First Difference'].append(p_value_diff)
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("ADF Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nADF Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))


# Save DataFrames as LaTeX tables
df_results_constant.to_latex('adf_test_constant.tex', index=False, caption='ADF Test with Constant Only')
df_results_trend.to_latex('adf_test_trend.tex', index=False, caption='ADF Test with Constant and Trend')
```

```{python}

import pandas as pd
import numpy as np
from arch.unitroot import ADF

# Function to perform the ADF test and return the t-statistic and p-value
def run_adf(ts, trend='c'):
    adf_test = ADF(ts.dropna(), trend=trend)
    return adf_test.stat, adf_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'



# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_adf(ts, trend='c')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(round(t_stat_level, 3))
    results_constant['p-value Level'].append(round(p_value_level, 3))
    results_constant['t-value First Difference'].append(round(t_stat_diff, 3))
    results_constant['p-value First Difference'].append(round(p_value_diff, 3))
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_adf(ts, trend='ct')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(round(t_stat_level, 3))
    results_trend['p-value Level'].append(round(p_value_level, 3))
    results_trend['t-value First Difference'].append(round(t_stat_diff, 3))
    results_trend['p-value First Difference'].append(round(p_value_diff, 3))
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Save DataFrames as LaTeX tables using Styler
df_results_constant.style.format(precision=3).to_latex('adf_test_constant.tex', caption='ADF Test with Constant Only', position_float='centering', index=False)
df_results_trend.style.format(precision=3).to_latex('adf_test_trend.tex', caption='ADF Test with Constant and Trend', position_float='centering', index=False)

# Printing the tables for verification
print("ADF Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nADF Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))

```


### 2.1.2 Phillips-Perron (PP)

```{python}


# Function to perform the Phillips-Perron test and return the test statistic and p-value
def run_pp(ts, trend='c'):
    pp_test = PhillipsPerron(ts.dropna(), trend=trend)
    return pp_test.stat, pp_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'

# Assuming ts_list is a dictionary with time series data as its values
# The keys of the dictionary are the names of the variables

# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_pp(ts, trend='c')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(t_stat_level)
    results_constant['p-value Level'].append(p_value_level)
    results_constant['t-value First Difference'].append(t_stat_diff)
    results_constant['p-value First Difference'].append(p_value_diff)
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_pp(ts, trend='ct')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(t_stat_level)
    results_trend['p-value Level'].append(p_value_level)
    results_trend['t-value First Difference'].append(t_stat_diff)
    results_trend['p-value First Difference'].append(p_value_diff)
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("Phillips-Perron Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nPhillips-Perron Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))


```


```{python}

import pandas as pd
import numpy as np
from arch.unitroot import PhillipsPerron

# Function to perform the Phillips-Perron test and return the test statistic and p-value
def run_pp(ts, trend='c'):
    pp_test = PhillipsPerron(ts.dropna(), trend=trend)
    return pp_test.stat, pp_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'


# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_pp(ts, trend='c')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(round(t_stat_level, 3))
    results_constant['p-value Level'].append(round(p_value_level, 3))
    results_constant['t-value First Difference'].append(round(t_stat_diff, 3))
    results_constant['p-value First Difference'].append(round(p_value_diff, 3))
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_pp(ts, trend='ct')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(round(t_stat_level, 3))
    results_trend['p-value Level'].append(round(p_value_level, 3))
    results_trend['t-value First Difference'].append(round(t_stat_diff, 3))
    results_trend['p-value First Difference'].append(round(p_value_diff, 3))
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("Phillips-Perron Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nPhillips-Perron Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))

# Save DataFrames as LaTeX tables using Styler
df_results_constant.style.format(precision=3).hide(axis='index').to_latex('pp_test_constant.tex', caption='Phillips-Perron Test with Constant Only', position_float='centering')
df_results_trend.style.format(precision=3).hide(axis='index').to_latex('pp_test_trend.tex', caption='Phillips-Perron Test with Constant and Trend', position_float='centering')

```



### 2.1.3 Order of Integration

```{python}

# Creating a list of dictionaries with each containing a key and a differenced series
ts_list_difff = {
    'carbon': ts_list['carbon'].diff().dropna(),
    'methane': ts_list['methane'].diff().diff().dropna(),
    'temperatures': ts_list['temperatures'].diff().dropna(),
    'ocean_warming': ts_list['ocean_warming'].diff().dropna(),
    'sea_level': ts_list['sea_level'].diff().dropna(),
    'drought': ts_list['drought'].diff().dropna(),  
    'flood': ts_list['flood'].diff().dropna(),
    'landslide': ts_list['landslide'].diff().dropna(),
    'storm': ts_list['storm'],  
    'wildfire': ts_list['wildfire'].diff().dropna(),
    'arctic_ice': ts_list['arctic_ice'].diff().diff().dropna(),
    'remittance': ts_list['remittence'].diff().dropna()}
    
    
    
```

```{python}


# Dictionary of variables and their respective order of integration
order_of_integration = {
    'carbon': 'I(1)',
    'methane': 'I(2)',
    'temperatures': 'I(1)',
    'ocean_warming': 'I(1)',
    'sea_level': 'I(1)',
    'drought': 'I(1)',
    'flood': 'I(1)',
    'landslide': 'I(1)',
    'storm': 'I(0)',  # Assuming 'storm' is stationary without differencing
    'wildfire': 'I(1)',
    'arctic_ice': 'I(2)',
    'remittance': 'I(1)'
}

# Create a DataFrame from the dictionary
integration_order_df = pd.DataFrame(list(order_of_integration.items()), columns=['Variable', 'Order of Integration'])

# To hide the index when printing
print(integration_order_df.to_string(index=False))

integration_order_df.to_latex('integration_order_table.tex', index=False, caption='Order of Integration for Variables', label='tab:integration_order')



```

```{python}

# Function to perform the ADF test and return the t-statistic and p-value
def run_adf(ts, trend=''):
    adf_test = ADF(ts.dropna(), trend=trend)
    return adf_test.stat, adf_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'


# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list_difff.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_adf(ts, trend='c')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(t_stat_level)
    results_constant['p-value Level'].append(p_value_level)
    results_constant['t-value First Difference'].append(t_stat_diff)
    results_constant['p-value First Difference'].append(p_value_diff)
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_adf(ts, trend='ct')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(t_stat_level)
    results_trend['p-value Level'].append(p_value_level)
    results_trend['t-value First Difference'].append(t_stat_diff)
    results_trend['p-value First Difference'].append(p_value_diff)
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("ADF Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nADF Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))


```

```{python}
import pandas as pd
from arch.unitroot import ADF

# Function to perform the ADF test and return the t-statistic and p-value
def run_adf(ts, trend=''):
    adf_test = ADF(ts.dropna(), trend=trend)
    return adf_test.stat, adf_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'


# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list_difff.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_adf(ts, trend='c')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(round(t_stat_level, 3))
    results_constant['p-value Level'].append(round(p_value_level, 3))
    results_constant['t-value First Difference'].append(round(t_stat_diff, 3))
    results_constant['p-value First Difference'].append(round(p_value_diff, 3))
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_adf(ts, trend='ct')
    t_stat_diff, p_value_diff = run_adf(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(round(t_stat_level, 3))
    results_trend['p-value Level'].append(round(p_value_level, 3))
    results_trend['t-value First Difference'].append(round(t_stat_diff, 3))
    results_trend['p-value First Difference'].append(round(p_value_diff, 3))
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("ADF Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nADF Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))

# Save DataFrames as LaTeX tables using Styler
df_results_constant.style.format(precision=3).hide(axis='index').to_latex('adf_test_constant_0.tex', caption='ADF Test with Constant Only', position_float='centering')
df_results_trend.style.format(precision=3).hide(axis='index').to_latex('adf_test_trend_0.tex', caption='ADF Test with Constant and Trend', position_float='centering')

```


```{python}
from arch.unitroot import PhillipsPerron

# Function to perform the Phillips-Perron test and return the test statistic and p-value
def run_pp(ts, trend='c'):
    pp_test = PhillipsPerron(ts.dropna(), trend=trend)
    return pp_test.stat, pp_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'

# Assuming ts_list is a dictionary with time series data as its values
# The keys of the dictionary are the names of the variables

# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list_difff.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_pp(ts, trend='c')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(t_stat_level)
    results_constant['p-value Level'].append(p_value_level)
    results_constant['t-value First Difference'].append(t_stat_diff)
    results_constant['p-value First Difference'].append(p_value_diff)
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_pp(ts, trend='ct')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(t_stat_level)
    results_trend['p-value Level'].append(p_value_level)
    results_trend['t-value First Difference'].append(t_stat_diff)
    results_trend['p-value First Difference'].append(p_value_diff)
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("Phillips-Perron Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nPhillips-Perron Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))
```


```{python}


import pandas as pd
from arch.unitroot import PhillipsPerron

# Function to perform the Phillips-Perron test and return the test statistic and p-value
def run_pp(ts, trend='c'):
    pp_test = PhillipsPerron(ts.dropna(), trend=trend)
    return pp_test.stat, pp_test.pvalue  # t-statistic, p-value

# Function to determine the order of integration based on p-values
def determine_order(p_value_level, p_value_diff):
    if p_value_level < 0.05:
        return 'I(0)'
    elif p_value_diff < 0.05:
        return 'I(1)'
    else:
        return 'I(>1)'


# DataFrames to store results
results_constant = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                    't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}
results_trend = {'Variable': [], 't-value Level': [], 'p-value Level': [],
                 't-value First Difference': [], 'p-value First Difference': [], 'Conclusion': []}

# Run tests for constant and constant with trend
for variable, ts in ts_list_difff.items():
    # Ensure that the time series is a pandas Series and has no missing values
    ts = pd.Series(ts).dropna()

    # Skip the loop if the series is empty after dropping NaN values
    if ts.empty:
        continue

    # Test with constant
    t_stat_level, p_value_level = run_pp(ts, trend='c')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='c')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_constant['Variable'].append(variable)
    results_constant['t-value Level'].append(round(t_stat_level, 3))
    results_constant['p-value Level'].append(round(p_value_level, 3))
    results_constant['t-value First Difference'].append(round(t_stat_diff, 3))
    results_constant['p-value First Difference'].append(round(p_value_diff, 3))
    results_constant['Conclusion'].append(conclusion)
    
    # Test with constant and trend
    t_stat_level, p_value_level = run_pp(ts, trend='ct')
    t_stat_diff, p_value_diff = run_pp(ts.diff().dropna(), trend='ct')
    conclusion = determine_order(p_value_level, p_value_diff)
    results_trend['Variable'].append(variable)
    results_trend['t-value Level'].append(round(t_stat_level, 3))
    results_trend['p-value Level'].append(round(p_value_level, 3))
    results_trend['t-value First Difference'].append(round(t_stat_diff, 3))
    results_trend['p-value First Difference'].append(round(p_value_diff, 3))
    results_trend['Conclusion'].append(conclusion)

# Convert the results to Pandas DataFrames
df_results_constant = pd.DataFrame(results_constant)
df_results_trend = pd.DataFrame(results_trend)

# Printing the tables
print("Phillips-Perron Test with Constant Only:")
print(df_results_constant.to_string(index=False))
print("\nPhillips-Perron Test with Constant and Trend:")
print(df_results_trend.to_string(index=False))

# Save DataFrames as LaTeX tables using Styler
df_results_constant.style.format(precision=3).hide(axis='index').to_latex('pp_test_constant_0.tex', caption='Phillips-Perron Test with Constant Only', position_float='centering')
df_results_trend.style.format(precision=3).hide(axis='index').to_latex('pp_test_trend_0.tex', caption='Phillips-Perron Test with Constant and Trend', position_float='centering')


```


Now all our variables are stationary

```{r}

ts_list_difff=list(
  #refugees = diff(ts_refugees, differences = 1),
  carbon = diff(ts_list[["carbon"]], differences = 1),
  methane = diff(ts_list[["methane"]], differences = 2),
  temperatures = diff(ts_list[["temperatures"]],differences = 1),
  ocean_warming = diff(ts_list[["ocean_warming"]],differences = 1),
  sea_level = diff(ts_list[["sea_level"]],differences = 1),
  drought = diff(ts_list[["drought"]],differences = 1),
  flood = diff(ts_list[["flood"]],differences = 1),
  landslide = diff(ts_list[["landslide"]],differences = 1),
  storm = ts_list[["storm"]],
  wildfire = diff(ts_list[["wildfire"]],differences = 1),
  arctic_ice = diff(ts_list[["arctic_ice"]],differences = 2),
  remittence=diff(ts_list[["remittence"]],differences = 1)
)

```

## 2.2 CORRELATIONS

```{r}

ts_matrix <- do.call(cbind, lapply(ts_list_difff, as.numeric))
cor_matrix <- cor(ts_matrix)
print(cor_matrix)


cor_matrix_1 <- cor_matrix[, 1:6]  # First 6 columns
cor_matrix_2 <- cor_matrix[, 7:12]  # Next 6 columns

# Convert each segment to a LaTeX table
cor_matrix_latex_1 <- xtable(cor_matrix_1, caption = "Correlation Matrix of Time Series (Part 1)")
cor_matrix_latex_2 <- xtable(cor_matrix_2, caption = "Correlation Matrix of Time Series (Part 2)")

# Save the LaTeX tables to .tex files
print(cor_matrix_latex_1, file = "correlation_matrix_1_diff.tex")
print(cor_matrix_latex_2, file = "correlation_matrix_2_diff.tex")

```

### 2.2.1 Multicollinearity

As we have differenciated the time series, now they have different length, so we will truncate it to the lowest length to be able to perform VIF

```{r}

shortest_length <- min(sapply(ts_list_difff, length))

# Now trim each series to have the same length
ts_list_difff <- lapply(ts_list_difff, function(x) tail(x, shortest_length))

# Combine the truncated time series into a data frame
df <- data.frame(ts_list_difff)

# Now you can run your VIF analysis
# Fit a linear model with all variables as predictors
lm_model <- lm(remittence ~ ., data = df) # Replace 'remittance' with the actual name of your dependent variable
vif_results <- vif(lm_model)
print(vif_results)



vif_df <- as.data.frame(vif_results)
vif_df$Variable <- rownames(vif_df)
rownames(vif_df) <- NULL

# Convert the VIF results to a LaTeX table
vif_latex <- xtable(vif_df, caption = "Variance Inflation Factor (VIF) Results")

# Save the LaTeX table to a .tex file
print(vif_latex, file = "vif_results_diff.tex", include.rownames = FALSE)

```


## SPLIT

```{r, eval=FALSE}

# Calculate the split point for 80/20 split
split_point <- round(length(ts_list_difff$carbon) * 0.85)

# Initialize empty lists for train and test sets
ts_list_train <- list()
ts_list_test <- list()

# Loop through each time series in the list and split them
for (var_name in names(ts_list_difff)) {
  ts_data <- ts_list_difff[[var_name]]
  
  # Be aware that ts_data must be a numeric vector for the following to work.
  # If it's not, you need to convert it beforehand.
  
  # Split the data into training and testing sets
  ts_list_train[[var_name]] <- ts_data[1:split_point]
  ts_list_test[[var_name]]  <- ts_data[(split_point+1):length(ts_data)]
}



```



```{python}

# Find the shortest length among all series
shortest_length = min(len(ts_data) for ts_data in ts_list_difff.values())

# Trim each series to have the same shortest length
ts_list_difff = {var_name: ts_data[-shortest_length:] for var_name, ts_data in ts_list_difff.items()}

```


```{python, eval=FALSE}
# Calculate the split point for 85/15 split
split_point = round(len(ts_list_difff['carbon']) * 0.85)

# Initialize empty dictionaries for train and test sets
ts_list_train = {}
ts_list_test = {}

# Loop through each time series in the dictionary and split them
for var_name, ts_data in ts_list_difff.items():
    # Split the data into training and testing sets
    ts_list_train[var_name] = ts_data[:split_point]
    ts_list_test[var_name] = ts_data[split_point:]

# Now ts_list_train and ts_list_test contain the training and testing data respectively


```



Now that the series are stationary we dont have multicollinearity.

## 2.2 VAR

### 2.2.1 Lag Order Selection

```{python}

from statsmodels.tsa.api import VAR

def select_var_lag_order(ts_list_difff):
    df = pd.DataFrame(ts_list_difff).dropna()
    model = VAR(df)
    max_lags = 1
    selected_orders = model.select_order(maxlags=max_lags)
    print(selected_orders.summary())
    
    # Return the selected order
    return selected_orders.selected_orders['aic']

print("Lag order selection for model:")
lag_order = select_var_lag_order(ts_list_difff)


```

The selected lag order is 1. We can not try with higher lags as with 2 or higher is too large for the number of observations and the number of equations. The largest model cannot be estimated.

### 2.2.2. VAR Model


#### Python

```{python}


model = VAR(pd.DataFrame(ts_list_difff).dropna())
var_results = model.fit(lag_order,trend='c')

print(var_results.summary())

with open('var_model_summary.txt', 'w') as f:
    f.write(var_results.summary().as_text())
```

#### 2.2.2.1 Stability of the model

The model is stable and normal

```{python}

# 1. Serial Correlation of Residuals
whiteness_test = var_results.test_whiteness()
print("Whiteness Test (Portmanteau) for Autocorrelation:")
print(whiteness_test.summary())

# 2. Stability of the VAR Model
# Check if the model is stable
stability=var_results.is_stable()
print("Stability Test:")
print(stability)


# 3. Residual Normality
normality_test_results = var_results.test_normality()
print(normality_test_results.summary())


```
```{python}
companion_matrix = var_results.coefs

# Flatten the companion matrix to obtain the eigenvalues
eigvals = np.linalg.eigvals(companion_matrix.reshape(-1, companion_matrix.shape[-1]))

# Check if all eigenvalues lie within the unit circle
stability_check = np.all(np.abs(eigvals) < 1)

# Print results
print("Eigenvalues of the companion matrix:", eigvals)
print("All eigenvalues inside unit circle:", stability_check)
```

```{python}

import numpy as np
import matplotlib.pyplot as plt

# Define the eigenvalues
eigenvalues = np.array([
    0.56431648+0.03742536j,  0.56431648-0.03742536j,
    -0.04518828+0.77148376j, -0.04518828-0.77148376j,
    -0.03554533+0.46501739j, -0.03554533-0.46501739j,
     0.12009307+0.j,         -0.81710736+0.j,
    -0.36496152+0.26851932j, -0.36496152-0.26851932j,
    -0.64627534+0.j,         -0.39845033+0.j
])

# Plot the eigenvalues on the complex plane
plt.figure(figsize=(8, 8))
plt.axhline(0, color='black', lw=0.5)
plt.axvline(0, color='black', lw=0.5)
unit_circle = plt.Circle((0, 0), 1, color='r', fill=False, linestyle='--')
plt.gca().add_artist(unit_circle)

# Plotting the eigenvalues
plt.scatter(eigenvalues.real, eigenvalues.imag, c='b', marker='o')
plt.title('Eigenvalues of the Companion Matrix')
plt.xlabel('Real Part')
plt.ylabel('Imaginary Part')
plt.grid(True)
plt.xlim(-1.5, 1.5)
plt.ylim(-1.5, 1.5)
plt.gca().set_aspect('equal', adjustable='box')

# Save the figure in the working directory
plt.savefig("eigenvalues_companion_matrix.png")
plt.show()


```


same result computing by hand:

```{python}

coefs = var_results.coefs

# Number of endogenous variables (K) and lag order (p)
K = coefs.shape[1]
p = coefs.shape[0]

# Construct the companion matrix
companion_matrix = np.zeros((K * p, K * p))

# Fill the top row blocks with the coefficient matrices
for i in range(p):
    companion_matrix[:K, i*K:(i+1)*K] = coefs[i]

# Fill the sub-diagonal with identity matrices
if p > 1:
    companion_matrix[K:, :-K] = np.eye(K * (p - 1))

# Compute the eigenvalues of the companion matrix
eigvals = np.linalg.eigvals(companion_matrix)

# Print results
print("Eigenvalues of the companion matrix:", eigvals)

# Check if all eigenvalues lie within the unit circle
stability_check = np.all(np.abs(eigvals) < 1)
print("All eigenvalues inside unit circle:", stability_check)

```



#### R


```{r}

library(vars)

train_matrix <- do.call(cbind, ts_list_difff)
var_results <- VARselect(na.omit(train_matrix), lag.max = 12, type = "const")
optimal_lags <- var_results$selection["AIC(n)"]
optimal_lags
```

no podemos implementar 2 porque peta

```{r}
library(vars)
var_model <- VAR(na.omit(train_matrix), p = 1)

# Mostrar el resumen del modelo
summary(var_model)

```



#### Stability of the model

```{r}
library(vars)

# Assuming you've already fitted a VAR model called var_model

# Test for serial correlation
autocorrelation_results <- serial.test(var_model)
print(autocorrelation_results)

# Stability test
#var model shows its stable

# Multivariate normality test
normality_results <- normality.test(var_model)
print(normality_results)

```



### 2.2.3 Granger Causality

#### Python

```{python,eval=FALSE}
from statsmodels.tsa.stattools import grangercausalitytests

#granger causality between each pair of variables
print("Granger Causality Tests:")
for variable in var_results.names:
    for caused_by in var_results.names:
        if variable != caused_by:
            print(f"\nTesting causality from {caused_by} to {variable}:")
            test_result = grangercausalitytests(var_results.model.data.orig_endog[[caused_by, variable]], maxlag=lag_order, verbose=True)



#this test uses fitted VAR model for checking causality
import statsmodels.api as sm

granger_results = {}
variables = var_results.model.endog_names  # This gets the variable names used in the VAR model

for i in range(len(variables)):
    for j in range(len(variables)):
        if i != j:
            test_result = var_results.test_causality(variables[i], variables[j], kind='f')
            granger_results[(variables[i], variables[j])] = test_result.summary()

# You can now print or examine 'granger_results' to see the Granger causality test outputs
for key, value in granger_results.items():
    print(f"Causality from {key[0]} to {key[1]}:\n{value}\n")


```

##### Granger pairwise

```{python}

from statsmodels.tsa.stattools import grangercausalitytests

print("Granger Causality Tests:")
for variable in var_results.names:
    for caused_by in var_results.names:
        if variable != caused_by:
            print(f"\nTesting causality from {caused_by} to {variable}:")
            # Perform the Granger causality test without verbose output
            test_results = grangercausalitytests(var_results.model.data.orig_endog[[caused_by, variable]], maxlag=lag_order, verbose=False)
            
            # Iterate through results for each lag
            for lag, result in test_results.items():
                f_test_statistic, f_p_value, f_degrees_of_freedom, f_degrees_of_freedom_resid = result[0]['ssr_ftest']
                
                # Check if the p-value is less than 0.10 and print relevant details
                if f_p_value < 0.10:
                    print(f"Lag {lag}:")
                    print(f"F-Statistic: {f_test_statistic:.4f}, p-value: {f_p_value:.4f}")
                    print(f"Degrees of Freedom: {f_degrees_of_freedom}, Degrees of Freedom Residual: {f_degrees_of_freedom_resid}")



```


##### Granger using fitted VAR:


```{python}

import statsmodels.api as sm

# Dictionary to store the Granger causality results
granger_results = {}
variables = var_results.model.endog_names  # Get the variable names from the VAR model

# Loop through each pair of variables
for i in range(len(variables)):
    for j in range(len(variables)):
        if i != j:
            # Perform the Granger causality test
            test_result = var_results.test_causality(variables[i], variables[j], kind='f')
            # Extract the p-value from the test result summary
            pvalue = test_result.pvalue
            # Save the result if the p-value is less than 0.10
            if pvalue < 0.10:
                granger_results[(variables[i], variables[j])] = test_result.summary()

# Print the results where the p-value was below 0.10
for key, value in granger_results.items():
    print(f"Causality from {key[0]} to {key[1]}:\n{value}\n")


```

```{python}

import statsmodels.api as sm

# Specify the variable of interest
target_variable = 'remittance'

# Dictionary to store the Granger causality results related to the target variable
granger_results = {}
variables = var_results.model.endog_names  # Get the variable names from the VAR model

# Loop through each pair of variables to find causality related to the target variable
for var in variables:
    if var != target_variable:
        # Test if 'var' Granger-causes 'remittance'
        test_result_to_remittance = var_results.test_causality(target_variable, var, kind='f')
        if test_result_to_remittance.pvalue < 0.10:
            granger_results[(target_variable, var)] = test_result_to_remittance.summary()

        # Test if 'remittance' Granger-causes 'var'
        test_result_from_remittance = var_results.test_causality(var, target_variable, kind='f')
        if test_result_from_remittance.pvalue < 0.10:
            granger_results[(var, target_variable)] = test_result_from_remittance.summary()

# Print the results related to the target variable where the p-value was below 0.10
for key, value in granger_results.items():
    print(f"Causality from {key[0]} to {key[1]}:\n{value}\n")


```

all results
```{python, eval=FALSE}
import statsmodels.api as sm

# Dictionary to store the Granger causality results
granger_results = {}
variables = var_results.model.endog_names  # Get the variable names from the VAR model

# Loop through each pair of variables
for i in range(len(variables)):
    for j in range(len(variables)):
        if i != j:
            # Perform the Granger causality test
            test_result = var_results.test_causality(variables[i], variables[j], kind='f')
            # Save the result summary
            granger_results[(variables[i], variables[j])] = test_result.summary()

# Print all the results
for key, value in granger_results.items():
    print(f"Causality from {key[0]} to {key[1]}:\n{value}\n")

```






#### R

ALL VARIABLES ARE STATIONARY NOW SO WE CAN IMPLEMENT THE VAR MODEL We split the data into 80-20 to compute the VAR and after check accuracy.



```{r}
df_caus=as.data.frame(df)

i = 1
for(i in 1:ncol(df))
{
  variable = colnames(df)[i]
  cat("\n-------",variable,"--------\n")
 print(causality(var_model,cause = variable)) 
}

#---------------------------------------------------------------




```


### 2.2.3 Impulse-Response Functions

#### Python

```{python}
irf=var_results.irf(periods=5)
irf.plot()
plt.savefig("irf.png")

plt.show() 


fevd=var_results.fevd(5)
fevd.summary()
fevd.plot()
plt.show() 

```




```{python}
import matplotlib.pyplot as plt

# Get the IRF object from VAR results
irf = var_results.irf(periods=5)

# Use the correct variable name 'remittance'
variable_name = 'remittance'
assert variable_name in var_results.names, f"{variable_name} is not in the list of variables."

# Find the index of 'remittance' in the variable list
remittance_index = var_results.names.index(variable_name)

# Get the number of variables
num_vars = len(var_results.names)

# Plot only the first half of the variables
print("Plotting first half of the variables with confidence intervals:")
for i in range(num_vars // 2):
    if i != remittance_index:
        fig = irf.plot(orth=False, impulse=variable_name, response=var_results.names[i])
        plt.show()

# Plot the second half of the variables
print("Plotting second half of the variables with confidence intervals:")
for i in range(num_vars // 2, num_vars):
    if i != remittance_index:
        fig = irf.plot(orth=False, impulse=variable_name, response=var_results.names[i])
        plt.show()

```


```{python}
import matplotlib.pyplot as plt
import numpy as np

# Get the IRF object from VAR results
irf = var_results.irf(periods=5)

# Use the correct variable name 'remittance'
variable_name = 'remittance'
assert variable_name in var_results.names, f"{variable_name} is not in the list of variables."

# Store the axes from the figures
axes_list = []

# Plot the IRFs for all the variables including remittance itself
print("Plotting variables with confidence intervals:")
for i in range(len(var_results.names)):
    fig = irf.plot(orth=False, impulse=variable_name, response=var_results.names[i])
    axes_list.append(fig.axes[0])  # Store the axis
    plt.close(fig)  # Close the individual figure to avoid displaying it

# Create a new figure with a 6x2 grid
fig, axes = plt.subplots(nrows=6, ncols=2, figsize=(15, 20))
fig.suptitle(f'IRF of {variable_name} with Confidence Intervals', fontsize=16)

# Flatten the axes array for easy iteration
axes = axes.flatten()

# Copy the axes from the individual plots to the new figure
for plot_idx, src_ax in enumerate(axes_list):
    dst_ax = axes[plot_idx]

    # Transfer all elements from src_ax to dst_ax
    for line in src_ax.get_lines():
        dst_ax.plot(line.get_xdata(), line.get_ydata(), label=line.get_label(), color=line.get_color())
    
    for collection in src_ax.collections:
        dst_ax.add_collection(collection)

    # Remove the zero line by setting it to a non-visible color
    for line in dst_ax.get_lines():
        if np.all(line.get_ydata() == 0):
            line.set_visible(False)

    dst_ax.set_title(src_ax.get_title())
    dst_ax.set_xlabel(src_ax.get_xlabel())
    dst_ax.set_ylabel(src_ax.get_ylabel())
    dst_ax.legend()
    dst_ax.grid(True)

# Adjust layout and show the plot
plt.tight_layout(rect=[0, 0.03, 1, 0.95])
plt.show()



```





```{python}

import matplotlib.pyplot as plt

# Get the IRF object from VAR results
irf = var_results.irf(periods=5)

# Use the correct variable name 'remittance'
variable_name = 'remittance'
assert variable_name in var_results.names, f"{variable_name} is not in the list of variables."

# Find the index of 'remittance' in the variable list
remittance_index = var_results.names.index(variable_name)

# Store the figures
figures = []

# Plot the IRFs for all the variables including remittance itself
print("Plotting variables with confidence intervals:")
for i in range(len(var_results.names)):
    fig = irf.plot(orth=False, impulse=variable_name, response=var_results.names[i])
    figures.append(fig)

# Create a new figure with a 6x2 grid
fig, axes = plt.subplots(nrows=6, ncols=2, figsize=(15, 20))
fig.suptitle(f'IRF of {variable_name} with Confidence Intervals', fontsize=16)

# Flatten the axes array for easy iteration
axes = axes.flatten()

# Extract the axes from the individual plots and copy them to the new figure
for plot_idx, fig in enumerate(figures):
    src_ax = fig.axes[0]  # Source axis from individual plot
    dst_ax = axes[plot_idx]  # Destination axis in the grid

    # Copy data from the source axis to the destination axis
    for line in src_ax.get_lines():
        dst_ax.plot(line.get_xdata(), line.get_ydata(), label=line.get_label())

    # Copy the fill_between data for confidence intervals
    for poly in src_ax.collections:
        paths = poly.get_paths()
        if len(paths) > 0:
            path = paths[0]
            verts = path.vertices
            dst_ax.fill_between(verts[:, 0], verts[:, 1], alpha=0.3, color=poly.get_facecolor()[0])

    dst_ax.set_title(src_ax.get_title())
    dst_ax.set_xlabel(src_ax.get_xlabel())
    dst_ax.set_ylabel(src_ax.get_ylabel())
    dst_ax.legend()
    dst_ax.grid(True)

# Adjust layout and show the plot
plt.tight_layout(rect=[0, 0.03, 1, 0.95])
plt.show()


```



```{python}

import pandas as pd
import matplotlib.pyplot as plt
import numpy as np
import seaborn as sns

# Provided FEVD data for remittance
fevd_remittance = {
    'Horizon': [0, 1, 2, 3, 4],
    'carbon': [0.112912, 0.136980, 0.137228, 0.139137, 0.138671],
    'methane': [0.174177, 0.149611, 0.141285, 0.139958, 0.138730],
    'temperatures': [0.066003, 0.036748, 0.035334, 0.034962, 0.034923],
    'ocean_warming': [0.011666, 0.033343, 0.062510, 0.063797, 0.063853],
    'sea_level': [0.018226, 0.037556, 0.046640, 0.046199, 0.047085],
    'drought': [0.002893, 0.001589, 0.003779, 0.005096, 0.007991],
    'flood': [0.041631, 0.061330, 0.057159, 0.056489, 0.056381],
    'landslide': [0.034032, 0.038359, 0.041697, 0.043049, 0.042688],
    'storm': [0.009154, 0.069903, 0.066830, 0.066454, 0.065906],
    'wildfire': [0.041489, 0.046195, 0.046475, 0.047636, 0.048202],
    'arctic_ice': [0.003299, 0.003861, 0.003623, 0.004640, 0.004621],
    'remittance': [0.484518, 0.384524, 0.357440, 0.352585, 0.350950]
}

# Convert the dictionary to a DataFrame
df_fevd_remittance = pd.DataFrame(fevd_remittance)

# Set up the color palette
colors = sns.color_palette("Paired", len(df_fevd_remittance.columns) - 1)

# Plotting the FEVD for remittance using bar plots
fig, ax = plt.subplots(figsize=(8, 5))

# Position of bars on x-axis
bar_width = 0.8
positions = range(len(df_fevd_remittance))

# Loop through each variable and create a bar for each horizon
bottom = np.zeros(len(df_fevd_remittance))
for i, column in enumerate(df_fevd_remittance.columns[1:]):
    ax.bar(positions, df_fevd_remittance[column], bottom=bottom, color=colors[i], label=column, width=bar_width)
    bottom += df_fevd_remittance[column]

ax.set_title('Forecast Error Variance Decomposition (FEVD) for Remittance')
ax.set_xlabel('Horizon')
ax.set_ylabel('Proportion of Variance')
ax.set_xticks(positions)
ax.set_xticklabels(df_fevd_remittance['Horizon'])
ax.legend(loc='upper right', bbox_to_anchor=(1.2, 1))

plt.tight_layout()
plt.savefig("fevd_remittances.png")
plt.show()



```




## 3. Predictions


### R

```{r}
library(Metrics)

# Extract fitted values from the VAR model
fitted_values <- fitted(var_model)
# Assuming the actual data is in the first column of train_matrix
actual_values <- train_matrix[, 12]  # Adjust this if your actual data column is different

# Calculate MSE using the Metrics package
mse_value <- mse(actual_values, fitted_values[, 12])  # Ensure the right variable is selected
print(paste("MSE of the VAR model is:", mse_value))


```


#### Ventana movil
```{r}
library(vars)
library(Metrics)
library(ggplot2)

# Configurar parámetros iniciales
window_size <- 25
total_length <- nrow(train_matrix)
steps_ahead <- 1

# Listas para almacenar predicciones y valores reales
predictions <- vector("numeric", total_length - window_size)
actual_values <- vector("numeric", total_length - window_size)

# Bucle para la validación cruzada con ventana móvil
for (i in 1:(total_length - window_size)) {
  # Seleccionar datos de entrenamiento
  train_data <- train_matrix[1:(window_size + i ), ]
  
  # Ajustar modelo VAR
  var_model <- VAR(train_data, p = 1, type = "const")
  
  # Predecir el siguiente paso
  forecast <- predict(var_model, n.ahead = steps_ahead)
  
  # Almacenar la predicción y el valor real
  predictions[i] <- forecast$fcst[[12]][, "fcst"]
  actual_values[i] <- train_matrix[window_size + i, 12]
}

# Calcular el MSE usando Metrics
mse_value <- mse(actual_values, predictions)

# Imprimir el MSE
print(paste("MSE of the forecasts is:", mse_value))

# Imprimir los valores predichos y reales
print("Predicted values:")
print(predictions)
print("Actual values:")
print(actual_values)

# Crear un data frame para graficar
plot_data <- data.frame(
  Time = (window_size + 1):(window_size + length(predictions)),
  Prediction = predictions,
  Actual = actual_values
)

# Graficar los resultados usando ggplot2
ggplot(plot_data, aes(x = Time)) +
  geom_line(aes(y = Prediction, colour = "Predicted"), linetype = "dashed") +
  geom_line(aes(y = Actual, colour = "Actual")) +
  labs(title = "Comparison of VAR Model Predictions vs Actual Data Remittance",
       x = "Time", y = "Value",
       colour = "Legend") +
  theme_minimal() +
  theme(legend.position = "bottom")

```



### Python

```{python}

import pandas as pd
import matplotlib.pyplot as plt
from sklearn.metrics import mean_squared_error
from statsmodels.tsa.api import VAR

# Assuming ts_list_difff is your differenced time series data
df = pd.DataFrame(ts_list_difff).dropna()

# Create a new date range starting from 1992
new_index = pd.date_range(start='1992', periods=len(df), freq='Y')
df.index = new_index

# Assuming 'var_results' is your fitted VAR model results object
# and 'df' is your DataFrame with the actual data

# Extract fitted values (predictions from the VAR model)
fitted_values = var_results.fittedvalues

# Extract the actual values for 'remittances' which is assumed to be the last column
actual_values = df.iloc[:, -1]

# Adjust the index of the fitted values to match the actual values
fitted_values.index = df.index[var_results.k_ar:]

# Ensure the indices of the fitted values match the actual values correctly
fitted_values_aligned = fitted_values.iloc[:, -1]
actual_values_aligned = actual_values.iloc[var_results.k_ar:]

# Convert the datetime indices to years
actual_values_aligned.index = actual_values_aligned.index.year
fitted_values_aligned.index = fitted_values_aligned.index.year

# Calculate MSE using the actual values and the predictions for the same variable
mse_value_diff = mean_squared_error(actual_values_aligned, fitted_values_aligned)
print(f"MSE of the VAR model for remittances is: {mse_value_diff}")

# Plot the actual vs. fitted values for the remittances series
plt.figure(figsize=(10, 5))
plt.plot(actual_values_aligned.index, actual_values_aligned, label='Original Diff Remittances', color='blue')
plt.plot(fitted_values_aligned.index, fitted_values_aligned, label='Fitted Diff Remittances', linestyle='--', color='orange')

plt.title('Original vs Fitted Values for Remittances')
plt.xlabel('Year')
plt.ylabel('Remittance')
plt.legend()
plt.grid(True)
plt.savefig("mse_in_sample.png")
plt.show()


```


```{python}

new_index = pd.date_range(start='1990', periods=len(data), freq='Y')

# Set this new date range as the index for your DataFrame
data.index = new_index
import pandas as pd
import matplotlib.pyplot as plt
from sklearn.metrics import mean_squared_error
from statsmodels.tsa.api import VAR

# Assuming 'var_results' is your fitted VAR model results object
# and 'data' is your DataFrame with the original actual data including the remittances column

# Extract fitted values (predictions from the VAR model)
fitted_values_diff = var_results.fittedvalues

# Integrate the differenced fitted values to match the original scale
initial_fitted_value = data.iloc[var_results.k_ar - 1, -1]  # k_ar is the number of lags in the VAR model
fitted_values_level = fitted_values_diff.cumsum() + initial_fitted_value

# Ensure the indices match for correct alignment
fitted_values_level.index = data.index[var_results.k_ar:]

# Extract the actual remittances values from the original data
actual_values = data.iloc[:, -1]

# Subset the actual values to start from the same index as fitted values
actual_values_subset = actual_values[fitted_values_level.index]

# Calculate MSE using the actual values and the integrated predictions for the same variable
mse_value_int = mean_squared_error(actual_values_subset, fitted_values_level.iloc[:, -1])
print(f"MSE of the VAR model for remittances is: {mse_value_int}")

# Plot the actual vs. integrated fitted values for the remittances series
plt.figure(figsize=(10, 5))
plt.plot(actual_values_subset.index, actual_values_subset, label='Original Remittance', color='blue')
plt.plot(fitted_values_level.index, fitted_values_level.iloc[:, -1], label='Fitted Remittance', linestyle='--', color='orange')

plt.title('Original vs Fitted Values for Remittances')
plt.xlabel('Year')
plt.ylabel('Remittance')
plt.legend()
plt.grid(True)
plt.show()

actual_values
fitted_values_level
actual_values_subset
```


```{python}
results_table = pd.DataFrame({
    'MSE Type': ['Integrated Series', 'Differenced Series'],
    'MSE Value': [mse_value_int, mse_value_diff]
})

# Display the table
print(results_table)


from sklearn.metrics import mean_squared_error
import numpy as np
rmse_value_int = np.sqrt(mse_value_int)
rmse_value_diff = np.sqrt(mse_value_diff)
#print(f"RMSE of the VAR model for remittances on the integrated series is: {rmse_value_int}")
print(f"RMSE of the VAR model for remittances on the differenced series is: {rmse_value_diff}")


from sklearn.metrics import r2_score
r2_value_int = r2_score(actual_values_aligned, fitted_values_aligned)
r2_value_diff = r2_score(actual_values_subset, fitted_values_level.iloc[:, -1])
#print(f"R² of the VAR model for remittances on the integrated series is: {r2_value_int}")
#print(f"R² of the VAR model for remittances on the differenced series is: {r2_value_diff}")


```


#### Ventana

not moving window, first iteration 25 años, second iteration 26 años....

```{python, warning=FALSE, message=FALSE}
import pandas as pd
import numpy as np
from statsmodels.tsa.api import VAR
from sklearn.metrics import mean_squared_error
import matplotlib.pyplot as plt
import matplotlib.dates as mdates

# Assuming df is your DataFrame
df = pd.DataFrame(ts_list_difff)  # Your data might already be loaded as shown in the # # Convert index to only include the year
df.index = pd.to_datetime(df.index).year

# Define initial parameters
window_size = 25  # smaller window for example purposes
total_length = len(df)
steps_ahead = 1

# Lists to store predictions and actual values
predictions = []
actual_values = []

# Rolling window validation
for i in range(window_size, total_length):
    # Select training data
    train_data = df.iloc[:i]

    # Fit VAR model
    model = VAR(train_data)
    var_model = model.fit(lag_order,trend='c')

    # Predict the next step
    forecast = var_model.forecast(train_data.values[-var_model.k_ar:], steps=steps_ahead)
    
    # Store the forecast and actual value
    predictions.append(forecast[0][11])  # Example for 'remittance' column
    actual_values.append(df.iloc[i, 11])

# Calculate MSE
mse_value = mean_squared_error(actual_values, predictions)
print(f"MSE of the forecasts is: {mse_value}")

# Print actual and predicted values
print("Predicted values:")
print(predictions)
print("Actual values:")
print(actual_values)

# Plotting the results
plt.figure(figsize=(10, 5))
plt.plot(df.index[window_size:], predictions, label='Predicted', linestyle='--')
plt.plot(df.index[window_size:], actual_values, label='Actual')

plt.title('Comparison of VAR Model Predictions vs Actual Data')
plt.xlabel('Year')
plt.ylabel('Value')

# Ensure x-axis only shows years (setting major ticks to show every year if needed)
plt.xticks(df.index[window_size:], df.index[window_size:].values, rotation=45)

plt.legend()
plt.grid(True)
plt.show()
```

moving window, siempre 25 años y se va moviendo:


```{python}

import pandas as pd
import numpy as np
from statsmodels.tsa.api import VAR
from sklearn.metrics import mean_squared_error
import matplotlib.pyplot as plt
import matplotlib.dates as mdates

# Assuming df is your DataFrame and it is already loaded
df = pd.DataFrame(ts_list_difff)
df.index = pd.to_datetime(df.index).year  # Convert index to only include the year

# Define initial parameters
window_size = 25  # Rolling window size
lag_order = 1  # Set lag order as 1
steps_ahead = 1  # Predicting one step ahead

# Lists to store predictions and actual values
predictions = []
actual_values = []
prediction_years = []

# Rolling window validation
for i in range(window_size, len(df) - steps_ahead + 1):
    # Select training data for the current window
    train_data = df.iloc[i - window_size:i]

    # Fit VAR model
    model = VAR(train_data)
    var_model = model.fit(lag_order)

    # Predict the next step
    forecast = var_model.forecast(train_data.values[-var_model.k_ar:], steps=steps_ahead)

    # Store the forecast and actual value
    predictions.append(forecast[0][11])  # Example for 'remittance' column
    actual_values.append(df.iloc[i, 11])
    prediction_years.append(df.index[i])

# Calculate MSE
mse_value = mean_squared_error(actual_values, predictions)
print(f"MSE of the forecasts is: {mse_value}")

# Calculate RMSE
rmse_value = np.sqrt(mse_value)
print(f"RMSE of the forecasts is: {rmse_value}")

# Print actual and predicted values
print("Predicted values:")
print(predictions)
print("Actual values:")
print(actual_values)

# Plotting the results
plt.figure(figsize=(10, 6))
plt.plot(prediction_years, predictions, label='Predicted Diff Remittances', linestyle='--', color='orange')
plt.plot(prediction_years, actual_values, label='Original Diff Remittances', color='blue')

plt.title('Comparison of VAR Model Predictions vs Original Data')
plt.xlabel('Year')
plt.ylabel('Value')

# Ensure x-axis only shows years
plt.xticks(ticks=prediction_years, labels=prediction_years, rotation=0)

plt.legend()
plt.grid(True)
plt.savefig("mse_window.png")
plt.show()
rmse_value = np.sqrt(mse_value)


```


```{python}
import pandas as pd
import numpy as np
from statsmodels.tsa.api import VAR
from sklearn.metrics import mean_squared_error
import matplotlib.pyplot as plt
import matplotlib.dates as mdates

# Assuming df is your DataFrame and it is already loaded
df = pd.DataFrame(ts_list_difff)
df.index = pd.to_datetime(df.index).year  # Convert index to only include the year

# Define initial parameters
window_size = 25  # Rolling window size
lag_order = 1  # Set lag order as 1
steps_ahead = 1  # Predicting one step ahead

# Lists to store predictions and actual values
predictions = []
actual_values = []
prediction_years = []

# Rolling window validation
for i in range(window_size, len(df) - steps_ahead + 1):
    # Select training data for the current window
    train_data = df.iloc[i - window_size:i]

    # Fit VAR model
    model = VAR(train_data)
    var_model = model.fit(lag_order)

    # Predict the next step
    forecast = var_model.forecast(train_data.values[-var_model.k_ar:], steps=steps_ahead)

    # Store the forecast and actual value
    predictions.append(forecast[0][11])  # Example for 'remittance' column
    actual_values.append(df.iloc[i, 11])
    prediction_years.append(df.index[i])

# Calculate MSE
mse_value = mean_squared_error(actual_values, predictions)
print(f"MSE of the forecasts is: {mse_value}")

# Calculate RMSE
rmse_value = np.sqrt(mse_value)
print(f"RMSE of the forecasts is: {rmse_value}")

# Forecasting the next 5 years
future_years = [year for year in range(df.index[-1] + 1, df.index[-1] + 6)]
last_window = df.iloc[-window_size:]  # Last available window of data

for year in future_years:
    # Fit VAR model on the last window
    model = VAR(last_window)
    var_model = model.fit(lag_order)

    # Predict the next step
    forecast = var_model.forecast(last_window.values[-var_model.k_ar:], steps=steps_ahead)

    # Append the forecast to the predictions and update the window
    last_window = last_window.append(pd.DataFrame([forecast[0]], columns=last_window.columns, index=[year]))
    predictions.append(forecast[0][11])
    prediction_years.append(year)

# Print actual and predicted values
print("Predicted values:")
print(predictions)
print("Actual values:")
print(actual_values)

# Plotting the results
plt.figure(figsize=(10, 6))
plt.plot(prediction_years, predictions, label='Predicted Diff Remittances', linestyle='--', color='orange')
plt.plot(df.index[-len(actual_values):], actual_values, label='Original Diff Remittances', color='blue')

plt.title('Comparison of VAR Model Forecast vs Original Data')
plt.xlabel('Year')
plt.ylabel('Value')

# Ensure x-axis only shows years
plt.xticks(ticks=prediction_years, labels=prediction_years, rotation=0)

plt.legend()
plt.grid(True)
plt.savefig("mse_window_forecast.png")
plt.show()

```

```{python}
import pandas as pd
import numpy as np
from statsmodels.tsa.api import VAR
import matplotlib.pyplot as plt
import matplotlib.dates as mdates

# Assuming df is your DataFrame and it is already loaded
df = pd.DataFrame(ts_list_difff)
df.index = pd.to_datetime(df.index).year  # Convert index to only include the year

# Define initial parameters
window_size = 25  # Rolling window size
lag_order = 1  # Set lag order as 1
steps_ahead = 1  # Predicting one step ahead

# Lists to store predictions and actual values
predictions_to_convert = df.iloc[:window_size, 11].tolist()  # Include initial values from the original data
actual_values = []
prediction_years = df.index[:window_size].tolist()

# Rolling window validation
for i in range(window_size, len(df) - steps_ahead + 1):
    # Select training data for the current window
    train_data = df.iloc[i - window_size:i]

    # Fit VAR model
    model = VAR(train_data)
    var_model = model.fit(lag_order)

    # Predict the next step
    forecast = var_model.forecast(train_data.values[-var_model.k_ar:], steps=steps_ahead)

    # Store the forecast and actual value
    predictions_to_convert.append(forecast[0][11])  # Example for 'remittance' column
    actual_values.append(df.iloc[i, 11])
    prediction_years.append(df.index[i])

# Forecasting the next 5 years
future_years = [year for year in range(df.index[-1] + 1, df.index[-1] + 6)]
last_window = df.iloc[-window_size:]  # Last available window of data

for year in future_years:
    # Fit VAR model on the last window
    model = VAR(last_window)
    var_model = model.fit(lag_order)

    # Predict the next step
    forecast = var_model.forecast(last_window.values[-var_model.k_ar:], steps=steps_ahead)

    # Append the forecast to the predictions and update the window
    last_window = last_window.append(pd.DataFrame([forecast[0]], columns=last_window.columns, index=[year]))
    predictions_to_convert.append(forecast[0][11])
    prediction_years.append(year)

# Print actual and predicted values
print("Predicted values to convert:")
print(predictions_to_convert)


```

```{python}
print(predictions_to_convert)

initial_value_1990 = ts_remittences.loc['1990']

# Value from the cut year 1991
initial_value_1991 = -0.06513

# Start the reconstruction process
predictions_converted_full = [initial_value_1990, initial_value_1990 + initial_value_1991]
for value in predictions_to_convert:
    predictions_converted_full.append(predictions_converted_full[-1] + value)

# Print the converted values
print("Predictions converted to original scale:")
print(predictions_converted_full)

# Create a date range starting from 1990 to match the original series
start_year = 1990
years = pd.date_range(start=f'{start_year}', periods=len(predictions_converted_full), freq='Y')

# Create a DataFrame for plotting
df_predictions = pd.DataFrame({'Year': years, 'Predicted Remittances': predictions_converted_full})
df_predictions.set_index('Year', inplace=True)

# Create a DataFrame for the original series with the same years
df_original = pd.DataFrame({'Year': years[:len(ts_remittences)], 'Original Remittances': ts_remittences})
df_original.set_index('Year', inplace=True)

# Plot the results
plt.figure(figsize=(10, 6))
plt.plot(df_original.index, df_original['Original Remittances'], linestyle='-', color='blue', label='Original Remittances')
plt.plot(df_predictions.index, df_predictions['Predicted Remittances'], linestyle='--', color='orange', label='Predicted Remittances')
plt.title('Comparison of VAR Model Forecast vs Original Data with Original Scale')
plt.xlabel('Year')
plt.ylabel('Remittances')
plt.legend()
plt.grid(True)
plt.show()

# Print the converted values with their corresponding years
print("Predictions converted to original scale with years:")
print(df_predictions)

plt.savefig("mse_window_forecast_original.png")

df_predictions_relevant = df_predictions.loc['2017':'2021']
df_predictions_relevant

df_original_relevant = df_original.loc['2017':'2021']
df_original_relevant


df_predictions_relevant = df_predictions_relevant.sort_index()
df_original_relevant = df_original_relevant.sort_index()

# Check if indices are aligned
if not df_predictions_relevant.index.equals(df_original_relevant.index):
    raise ValueError("The indices of the prediction and original data do not match.")

# Calculate MSE and RMSE for the relevant period
mse = mean_squared_error(df_original_relevant, df_predictions_relevant)
rmse = np.sqrt(mse)

print(f'MSE: {mse}')
print(f'RMSE: {rmse}')

```



## 4. NNVAR



```{python}
import numpy as np
from tensorflow.keras.models import Sequential
from tensorflow.keras.layers import LSTM
from tensorflow.keras.layers import Dense, Dropout
import pandas as pd
from matplotlib import pyplot as plt
from sklearn.preprocessing import StandardScaler
import seaborn as sns
#from datetime import datetime

df=pd.DataFrame(ts_list)

train_dates = df.index
print(train_dates[-15:])  # Check last few dates

# Variables for training
cols = df.columns
print(cols) # Print columns to verify

df_for_training = df.astype(float)

# Normalize the dataset
scaler = StandardScaler()
scaler = scaler.fit(df_for_training)
df_for_training_scaled = scaler.transform(df_for_training)


#df_for_training_scaled = pd.DataFrame(df_for_training_scaled, columns=df_for_training.columns, index=df.index)
#print(df_for_training_scaled.head()) 

n_past = 14   # Number of past days we want to use to predict the future
n_predict = 1  # Number of future days we want to predict

trainX = []
trainY = []

#Reformat input data into a shape: (n_samples x timesteps x n_features)
#In my example, my df_for_training_scaled has a shape (12823, 5)
#12823 refers to the number of data points and 5 refers to the columns (multi-variables).
for i in range(n_past, len(df_for_training_scaled) - n_future +1):
    trainX.append(df_for_training_scaled[i - n_past:i, 0:df_for_training.shape[1]])
    trainY.append(df_for_training_scaled[i + n_future - 1:i + n_future, 0])


trainX, trainY = np.array(trainX), np.array(trainY)

print('trainX shape == {}.'.format(trainX.shape))
print('trainY shape == {}.'.format(trainY.shape))


# Define the LSTM model
model = Sequential()
model.add(LSTM(64, activation='relu', input_shape=(trainX.shape[1], trainX.shape[2]), return_sequences=True))
model.add(LSTM(32, activation='relu', return_sequences=False))
model.add(Dropout(0.2))
model.add(Dense(1))

model.compile(optimizer='adam', loss='mse')
model.summary()

# Fit the model
history = model.fit(trainX, trainY, epochs=50, batch_size=16, validation_split=0.1, verbose=1)

print(history.history.keys())

# Print the loss values for each epoch
print("Training loss values: ", history.history['loss'])
print("Validation loss values: ", history.history['val_loss'])

# Plot the training and validation loss values
plt.figure(figsize=(12, 6))
plt.plot(history.history['loss'], label='Training loss')
plt.plot(history.history['val_loss'], label='Validation loss')
plt.xlabel('Epochs')
plt.ylabel('Loss')
plt.title('Training and Validation Loss')
plt.legend()
plt.show()



n_future = 7  # Number of years we want to predict into the future
# Adjust to start forecasting from the year after the last available year
datelist_future = pd.date_range(datelist_train[-1], periods=n_future+1, freq='Y').tolist()[1:]
datelist_train = df.index

# Convert Pandas Timestamp to Datetime object (for transformation) --> FUTURE
datelist_future_ = [this_timestamp.date() for this_timestamp in datelist_future]

# Make predictions for future dates
predictions_future = model.predict(trainX[-1:])


prediction_copies = np.repeat(predictions_future, df_for_training.shape[1], axis=-1)
y_pred_future = scaler.inverse_transform(prediction_copies)[:,-1]
datelist_train = df.index

# Create DataFrame for future predictions
PREDICTIONS_FUTURE = pd.DataFrame(y_pred_future, columns=['Predicted Remittences']).set_index(pd.Series(datelist_future_))

# Plotting
plt.figure(figsize=(14, 8))
sns.lineplot(data=df, x=df.index, y='remittence', label='Original Remittences')
sns.lineplot(data=PREDICTIONS_FUTURE, x=PREDICTIONS_FUTURE.index, y='Predicted Remittences', label='Predicted Remittences')
plt.axvline(x=df.index[-1], color='blue', linestyle='--')
plt.legend()
plt.show()

```






```{python}
import numpy as np
import pandas as pd
import random
import tensorflow as tf
from sklearn.preprocessing import MinMaxScaler
from sklearn.metrics import mean_squared_error
import matplotlib.pyplot as plt
from tensorflow.keras.models import Sequential
from tensorflow.keras.layers import LSTM, Dense, Input

# Set the seed for reproducibility
def set_seed(seed):
    np.random.seed(seed)
    random.seed(seed)
    tf.random.set_seed(seed)

set_seed(1)

# Load the data
remittence = df['remittence'].values

# Scale the data
scaler = MinMaxScaler(feature_range=(0, 1))
scaled_data = scaler.fit_transform(df)

# Create dataset for supervised learning
def create_dataset(dataset, time_step=1):
    dataX, dataY = [], []
    for i in range(len(dataset) - time_step - 1):
        dataX.append(dataset[i:(i + time_step), :])
        dataY.append(dataset[i + time_step, :])
    return np.array(dataX), np.array(dataY)

time_step = 10
X, y = create_dataset(scaled_data, time_step)
X = X.reshape(X.shape[0], X.shape[1], X.shape[2])  # [samples, time steps, features]

session_conf = tf.compat.v1.ConfigProto(intra_op_parallelism_threads=1, inter_op_parallelism_threads=1)
sess = tf.compat.v1.Session(graph=tf.compat.v1.get_default_graph(), config=session_conf)
tf.compat.v1.keras.backend.set_session(sess)
# Build the LSTM model
model = Sequential()
model.add(Input(shape=(time_step, df.shape[1])))
model.add(LSTM(50, return_sequences=True))
model.add(LSTM(50, return_sequences=False))
model.add(Dense(25))
model.add(Dense(df.shape[1]))  # Predict all features

model.compile(optimizer='adam', loss='mean_squared_error')
model.fit(X, y, epochs=50, batch_size=32, validation_split=0.2)

# Make predictions
predictions_scaled = model.predict(X)
predictions = scaler.inverse_transform(predictions_scaled)  # Inverse transform all features

# Calculate MSE and RMSE for the "remittence" column
mse = mean_squared_error(remittence[time_step+1:], predictions[:, df.columns.get_loc('remittence')])
rmse = np.sqrt(mse)
print(f'MSE: {mse}')
print(f'RMSE: {rmse}')

# Forecast the next 5 years
last_sequence = scaled_data[-time_step:, :].reshape((1, time_step, df.shape[1]))
future_predictions = []

for _ in range(5):
    next_pred = model.predict(last_sequence)
    future_predictions.append(next_pred[0, :])
    last_sequence = np.append(last_sequence[:, 1:, :], np.expand_dims(next_pred, axis=1), axis=1)

future_predictions = np.array(future_predictions)
future_predictions_inversed = scaler.inverse_transform(future_predictions)

# Extract the "remittence" column from future predictions
future_remittence = future_predictions_inversed[:, df.columns.get_loc('remittence')]

# Create extended remittence array
extended_remittence = np.concatenate((remittence, future_remittence))

data = np.array(extended_remittence)

# Years from 1990
years = np.arange(1990, 1990 + len(data))

# Plotting the data
plt.figure(figsize=(12, 6))
plt.plot(years, data, label=' Original Remittances', color='blue')
plt.plot(years[-5:], data[-5:], label='Predicted Remittances ', color='orange')
plt.xlabel('Year')
plt.ylabel('Remittence')
plt.title('Remittence Over Years with Values forecasted with LSTM')
plt.legend()
plt.grid(True)  # Añadir esta línea para mostrar la cuadrícula
plt.show()

#plt.savefig("mse_window_forecast_nn.png")


```


